---
title: "Accessing the Web App"
description: "Accessing the Web App"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accessing the Web App}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r developernote, eval=FALSE, echo= FALSE, include = FALSE}
#  *>>>>>>>>>> Developer note: vignettes need to be tested/edited/rebuilt regularly <<<<<<<<<<<*
#    - **See ?pkgdown::build_site** and script in EJAM/data-raw/- EJAM uses the pkgdown R package to build help and articles/ vignettes as web pages
```

```{r SETUP_default_eval_or_not, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(eval = FALSE)
# https://r-pkgs.org/vignettes.html
```

# Web app for the public via EJScreen's Multisite Tool

EPA has integrated EJAM's toolkit as the EJScreen Multisite Tool, in the EJScreen "Tools" menu. EPA's [EJScreen](https://ejscreen.epa.gov/mapper/ "https://ejscreen.epa.gov/mapper/"){.uri target="_blank" rel="noreferrer noopener"} provides public access to batches of EJ reports, with rapid multisite analysis, by leveraging the EJAM toolkit. This version is specifically configured for general use, so it does not include some less-frequently-used, complex features. Those needing the full set of complex analytic features can find them in the open source R package described below.

# Web app for EPA staff

EPA staff while on the Agency network can access an [internal version of the EJAM web app](https://usepa.sharepoint.com/sites/EJAMWorkgroup/SitePages/access-ejam.aspx){.uri target="_blank" rel="noreferrer noopener"}. It is specifically configured just for use by staff while at EPA offices or on EPA's VPN, and requires SSO authentication with a PIV card. *This is the same as the version any developer can run locally using the public code repository described below.*

# Web app for analysts or developers using R/RStudio

The EJAM software and data are available as open source resources, so that anyone using [R/RStudio](https://posit.co/download/rstudio-desktop/){.uri target="_blank" rel="noreferrer noopener"} (not just EPA staff) can use EJAM on their own computer.
 
Analysts or developers using R/RStudio have the option of running a local copy of the EJAM web app on their own computer. This may be even faster than relying on the hosted web app, does not time out after inactivity, and could be customized by a developer. You can also launch it with customized options or use bookmarked settings (and/or use EJAM functions and data directly without the web app, for more complex work).

You can install the EJAM R package and datasets as explained in [Installing the EJAM R package](1_installing.html). There is also a [Quick Start guide](2_quickstart.html) and extensive [documentation of EJAM functions/tools/data](../reference/index.html).

Once EJAM is installed, you can launch the local web app from RStudio as follows:

```{r run_app1, eval = FALSE, include= TRUE}
library(EJAM)
options(shiny.launch.browser = TRUE) # so the web app uses a browser (not the RStudio viewer)

run_app()
```

The EJScreen Multisite Tool configuration is available via `run_app(isPublic=T)` and the version for staff use is available via `run_app(isPublic=F)`

See documentation of optional parameters via `?run_app()`

Note this is different than running a simple shiny app that is not also a package via the golem package. You should use the EJAM package function `run_app()` not `shiny::runApp()`.
